---
title: "Voice pleasantness"
author: "Hille"
date: "March 14, 2017"
output: 
  html_document:
    dev: png
    fig_height: 6
---

```{r prepare, include=FALSE}
require(pryr)
require(plyr)
require(GGally)
require(ggplot2)
require(RColorBrewer)
require(psych)

options("width"=200)

```
## Data

Read the speaker scores: 

*  __speaker.gender__ speaker gender: __F__ female,  __M__ male
*  __speaker__ speaker voice sample nr in the test
*  __rater__ rater id
*  __gender__ rater gender
*  __age__ rater age
*  __score__ rater score

```{r}

scores = read.csv("scores.csv")
scores.f=scores[scores$speaker.gender=="F",]
scores.m=scores[scores$speaker.gender=="M",]

```

Read the GeMAPS parameters of the voice samples:

*  __speaker.gender__ speaker gender: __F__ female,  __M__ male
*  __speaker__ speaker voice sample nr in the test
*  __speaker.age__ speaker age
*  __phonogenre__
*  __GeMAPS__ the GeMAPS parameters

```{r}

voices = read.csv("voices.csv")
gemaps.start=5

```

## Intraclass Correlation Coefficient

Calculate the ICC for the different rater groups.

Select the female and male voices separately and create the data frames where the scores for each voice are in one row.
```{r ICC_1}

ids = c("speaker.gender","rater","gender","age")
xf = reshape(scores.f, idvar=ids, timevar = "speaker", direction="wide")
xm = reshape(scores.m, idvar=ids, timevar = "speaker", direction="wide")

```

Calculate __Intraclass Correlation Coefficient__ (ICC) for different groups:
```{r ICC_2}

subsets = c( 'T',
  'age=="Y"', 'age=="A"', 'gender=="M"','gender=="F"',
  'gender=="M" & age=="Y"', 'gender=="M" & age=="A"',
  'gender=="F" & age=="Y"','gender=="F" & age=="A"'
)

icc2k = function(data, sub ) {
  i = ICC(
    t(subset(data, subset=eval(parse(text=sub)),select=c(-1:-4))),
    missing = F)$results[5,]
  data.frame(subset=sub, ICC=i$ICC, F=i$F, p=i$p)
}
```
### Female speakers
```{r}
print(ldply(subsets, icc2k, data=xf))
```
### Male speakers
```{r}
print(ldply(subsets, icc2k, data=xm))

```



## Speaker score for the `phonogenre` groups

Print the boxplot of the __score__ grouped by __phonogenre__. 
```{r phonogenre_mean}

x = merge( scores, voices[c("speaker.gender","speaker","phonogenre")], by=c("speaker.gender","speaker"))

print(ggplot(x, aes(x=phonogenre, y=score, fill=phonogenre)) +
  geom_boxplot(notch=TRUE, width=.5) + 
  guides(fill=FALSE)+
  theme(axis.text = element_text(size=rel(1.5)))+
  theme(axis.title=element_text(size=rel(1.5), face="bold", colour="black"))
)

print( oneway.test(score ~ phonogenre, data=x))
```



## Two sample t.test for the female and male raters

Compare the male and female rater group scores for the female and male speakers with the t.test

```{r female M / F}

ttest = function( speaker.in, data)  {
  tt = t.test(
    subset(data, gender=="M" & speaker==speaker.in, select=c("score")),
    subset(data, gender=="F" & speaker==speaker.in, select=c("score"))
  )
  data.frame(statistic = tt$statistic, p.value = tt$p.value, parameter = tt$parameter)
}
print(ldply(1:50, ttest, data=scores.f))
print(ldply(1:60, ttest, data=scores.m))

```

## Two sample t.test for the male and female speakers

Calculate the mean score of the female and male voice samples for each rater 
and compare the female and male speaker scores with the paired
and the independent __two sample t.test__.
```{r}
x = aggregate(score ~ speaker.gender+rater, scores, FUN=mean)

x1 = merge(x[x$speaker.gender=="F",], x[x$speaker.gender=="M",], by="rater" )
names(x1) = c("rater","speaker.female","score.female","speaker.male","score.male")
t.test( x1$score.female, x1$score.male, paired=TRUE)

t.test( x1$score.female, x1$score.male)
```
### Boxplot
```{r}
print(ggplot(x, aes(x=speaker.gender, y=score, fill=speaker.gender)) +
  geom_boxplot(notch=TRUE, width=.33) + 
  xlab("speaker gender")+
  guides(fill=FALSE)+
  scale_x_discrete(labels=c("female speaker", "male speaker"))+
  theme(axis.text = element_text(size=rel(1.5)))+
  theme(axis.title = element_text(size=rel(1.5), face="bold", colour="black"))
)
```

## Speaker age correlation to the score

```{r}

x = merge(scores,voices[c("speaker.gender","speaker","speaker.age")],by=c("speaker.gender","speaker"))
x.f = x[x$speaker.gender=="F",]
x.m = x[x$speaker.gender=="M",]

cor.test(x$score, x$speaker.age)
cor.test(x.f$score, x.f$speaker.age)
cor.test(x.m$score, x.m$speaker.age)
```

### Plot the average score for all the speaker age groups

```{r}
x = aggregate(score ~ speaker.gender + speaker.age, data=x, FUN = mean )
print(
  ggplot(x,aes(x=speaker.age, y=score, color=speaker.gender))+
  geom_point(size=3)+
  stat_smooth(method = lm)+
  xlab("speaker age")+
  guides(color=guide_legend(title="Speaker"))+
  scale_colour_discrete(labels=c("female", "male"))+
  theme(legend.position = c(1,1), 
        legend.justification = c(1,1), 
        legend.background = element_blank(), 
        legend.key = element_blank())+
  theme(legend.title=element_text(size=rel(1.5), face="bold"))+
  theme(legend.text=element_text(size=rel(1.5)))+
  theme(axis.text = element_text(size=rel(1.5)))+
  theme(axis.title=element_text(size=rel(1.5), face="bold", colour="black"))
)

```

## Rater age and gender effect on the score

Calculate the average score of all the voice samples for all the groups and plot them.

```{r aggregate_score}

da = aggregate(score ~ speaker.gender+speaker+gender+age, data=scores, mean)
```


```{r compare_female, fig.asp=1}
dvf = da[da$speaker.gender=="F",]

mvft = matrix(dvf[ order(dvf$gender, dvf$age),]$score,50,4,byrow=F, 
      dimnames=list(1:50,c("adult_women", "young_women", "adult_men", "young_men")))

print(
  ggpairs(
    data.frame(mvft),title="Female voices", 
    columnLabels=c("women over 35", "women under 35", "men over 35","men under 35"),
    lower=list(continuous=wrap("points", colour="red" )),
    upper =list(continuous = wrap("cor", colour = "black"))
  )+
  theme(plot.title=element_text(size=rel(1.5)))+
  #theme(axis.text = element_text(size=rel(1.5)))+
  theme(axis.title=element_text(size=rel(1.5), face="bold", colour="black"))
)
```


```{r compare_male, fig.asp=1}
dvm = da[da$speaker.gender=="M",]

mvmt = matrix(dvm[order(dvm$gender, dvm$age),]$score,60,4,byrow=F, 
      dimnames=list(1:60,c("adult_women", "young_women", "adult_men", "young_men")))

print(
  ggpairs(
    data.frame(mvmt),title="Male voices",
    columnLabels=c("women over 35", "women under 35", "men over 35","men under 35"),
    lower=list(continuous=wrap("points", colour="blue" )),
    upper =list(continuous = wrap("cor", colour = "black"))
  )+
  theme(plot.title=element_text(size=rel(1.5)))+
  theme(axis.title=element_text(size=rel(1.5), face="bold", colour="black"))
)
```


## Plot the group average scores

```{r compare_lines, fig.width=15}
blue=brewer.pal(name="Blues",n=3)
red=brewer.pal(name="Reds",n=3)

print(
  ggplot(dvf,aes(x=speaker,y=score,color=paste(gender,age)))+
  geom_line(size=1.5,alpha=.3)+
  geom_point(size=3)+
  ylim(1,7)+
  ggtitle("Female voices")+
  labs(color="Groups")+
  xlab("voice number")+
  scale_y_continuous(breaks=c(2,3,4,5,6))+
  annotate("text", x=40, y=1.5, 
      label="Group average scores", size=8, 
      family="Helvetica", fontface="bold", colour=red[3])+
  scale_colour_manual(values=c(red[3],red[2],blue[3],blue[2]), 
      labels=c("women over 35", "women under 35", "men over 35","men under 35"))+
  theme(plot.title=element_text(size=rel(3), face="bold", colour="black"))+
  theme(legend.title=element_text(size=rel(2), face="bold"))+
  theme(legend.text=element_text(size=rel(1.5), face="bold"))+
  theme(axis.title=element_text(size=rel(1.5), face="bold", colour="black"))
)
```

```{r fig.width=15}
print(
  ggplot(dvm,aes(x=speaker,y=score,color=paste(gender,age)))+
  geom_line(size=1.5,alpha=.3)+
  geom_point(size=3)+
  ylim(1,7)+
  ggtitle("Male voices")+
  labs(color="Groups")+
  xlab("voice number")+
  scale_y_continuous(breaks=c(2,3,4,5,6))+
  annotate("text", x=50, y=1.5, 
      label="Group average scores", size=8, 
      family="Helvetica", fontface="bold", colour=blue[3])+
  scale_colour_manual(values=c(red[3],red[2],blue[3],blue[2]),
      labels=c("women over 35", "women under 35", "men over 35","men under 35"))+
  theme(plot.title=element_text(size=rel(3), face="bold", colour="black"))+
  theme(legend.title=element_text(size=rel(2), face="bold"))+
  theme(legend.text=element_text(size=rel(1.5), face="bold"))+
  theme(axis.title=element_text(size=rel(1.5), face="bold", colour="black"))
) 
```


## GeMAPS oneway.test (ANOVA)

Run the __oneway.test__ (Welch test) for all the GeMAPS parameters with `phonogenre` as the grouping factor. All the features are normalized.
The test is run separately for the female and male voices.

```{r anova}

l = length(names(voices))
vars = names(voices)[gemaps.start:l]

oneway = function(var, data) {
  x = oneway.test(formula(paste(var," ~ phonogenre")), data=data)
  data.frame(
    name=strsplit(x$data.name," ")[[1]][1], 
    F=x$statistic, 
    p.value=x$p.value,
    stringsAsFactors = F)
}

```
### Female voices
```{r}

x = voices[voices$speaker.gender=="F",]
voices.f = cbind( x[1:(gemaps.start-1)], scale(x[gemaps.start:l])[,] )
oneway.f = ldply(vars, oneway, data=voices.f)
oneway.f = oneway.f[ order(oneway.f$F, decreasing = TRUE),]
print(oneway.f)
```
### Male voices
```{r}

x = voices[voices$speaker.gender=="M",]
voices.m = cbind( x[1:(gemaps.start-1)], scale(x[gemaps.start:l])[,] )
oneway.m = ldply(vars, oneway, data=voices.m)
oneway.m = oneway.m[ order(oneway.m$F, decreasing = TRUE),]
print(oneway.m)

```


## GeMAPS t.test 
Run the __one sample t.test__ for all the GeMAPS parameters to get the CI interval.
The test is run separately for the female and male voices.

```{r plot_gemaps_female/male}

ttest = function(genre, data, name) {
  tt = t.test( (data[data$phonogenre==genre,])[[name]])
  data.frame(genre=genre,
      t=tt$statistic, 
      p.value=tt$p.value, 
      mean=tt$estimate, 
      ci.min=tt$conf.int[1], 
      ci.max=tt$conf.int[2])
}

smdensity = function(name, data) {
    #sm.density.compare(data[[name]], data$phonogenre, model="equal", xlab=name)
    #legend("topright", levels(data$phonogenre), fill=2+(0:nlevels(data$phonogenre)))
    cat("\n==== ", name, "\n")
    print(ldply( levels(voices$phonogenre), ttest, data=data, name=name ))
}
```
### Female voices
```{r}
l_ply(oneway.f$name, smdensity, data=voices.f)
```
### Male voices
```{r}
l_ply(oneway.m$name, smdensity, data=voices.m)

```

